### 2016-07-06 case-studies.tex
Reverted changes to ansi-isolation.

% To understand whether a straightforward implementation of the view
% counter preserves the invariant, we considered a \txnimp program
% that is representative of the counter application. In particular, we
% considered four concurrent transactions that operate on a shared
% counter: a reader transaction that reads the counter twice obtaining
% the values $k_1$ and $k_2$, respectively, and three writer
% transactions, each incrementing the counter once. If the reader
% witnesses a monotonically increasing view count, then $k_2\ge k_1$.
% We then instantiated our reasoning framework for an {\sc sc} store
% with {\sc rc} as the default isolation level (Postgres is one such
% store), with the aim of verifying the program under this
% instantiation.  Our attempts were however unsuccessful owing to our
% inability to prove that a writer transaction always commits a value
% no smaller than the current value of the counter. This points to the
% existence of anamolous executions where  To understand the
% prevalence of such anomalies, we implemented a view counter
% application in Ruby-on-Rails using standard templates and
% techniques.   
To understand the effect of changing store consistency levels, we
re-implemented the counter as an op-based \emph{replicated data type}
(RDT)~\cite{crdt,burckhardt14} on top of Cassandra
\emph{bolted-on}~\cite{bailisbolton} with {\sc cc} and {\sc sc}
consistency levels (both on-demand; {\sc ec} is default), and ANSI SQL
isolation levels. Keeping the consistency level at {\sc ec}, we then
repeated our experiments with committed reads and serializable
increments - a combination which was proved to preserve monotonicity
on an {\sc sc} store, but nonetheless observed monotonicity
violations. The reason for such violations became evident when
we instantiated our proof framework for an {\sc ec} store.
Under this instantiation, it is impossible to derive the
\C{MonotonicVis} (\S\ref{sec:ansi-isolation}) property needed to show
that successive reads in a session witness monotonically increasing
state. However, assuming {\sc cc} store consistency for reads allowed
us to complete the proof, and recover the monotonicity invariant. In
fact, we were able to complete the proof even after relaxing the
serializability requirement for increments, just by changing the
definition of the $\interp{\E}(X)$ (\S\ref{sec:opsem}) to return the
value of the largest write-to-$X$ instead of the latest write-to-$X$.
In other words, if the reads chose the largest valued write effect,
instead of the latest write effect they witness, then counter
increments need not be serializable for reads to witness a
monotonically increasing count.\footnote{The trade-off is that the view
count is now no longer accurate, although it grows monotonically.}
We implemented this version of the counter as an op-based RDT, which
uses {\sc cc} reads and {\sc ec} increments, both un-isolated (i.e.,
\emph{Read Uncommitted}), and obtained a 52\% improvement in overall
latency compared to the version that uses serializable increments.
\vspace*{-3pt}

>>>>

% To understand whether a straightforward implementation of the view
% counter preserves the invariant, we considered a \txnimp program
% that is representative of the counter application. In particular, we
% considered four concurrent transactions that operate on a shared
% counter: a reader transaction that reads the counter twice obtaining
% the values $k_1$ and $k_2$, respectively, and three writer
% transactions, each incrementing the counter once. If the reader
% witnesses a monotonically increasing view count, then $k_2\ge k_1$.
% We then instantiated our reasoning framework for an {\sc sc} store
% with {\sc rc} as the default isolation level (Postgres is one such
% store), with the aim of verifying the program under this
% instantiation.  Our attempts were however unsuccessful owing to our
% inability to prove that a writer transaction always commits a value
% no smaller than the current value of the counter. This points to the
% existence of anamolous executions where  To understand the
% prevalence of such anomalies, we implemented a view counter
% application in Ruby-on-Rails using standard templates and
% techniques.   
To understand the effect of changing store consistency levels, we
re-implemented the counter as an op-based \emph{replicated data type}
(RDT)~\cite{crdt,burckhardt14} on top of Cassandra
\emph{bolted-on}~\cite{bailisbolton} with {\sc cc} and {\sc sc}
consistency levels (both on-demand; {\sc ec} is default), and ANSI SQL
isolation levels. Keeping the consistency level at {\sc ec}, we then
repeated our experiments with committed reads and serializable
increments - a combination which was proved to preserve monotonicity
on an {\sc sc} store, but nonetheless observed monotonicity
violations. The reason for such violations became evident when
we instantiated our proof framework for an {\sc ec} store.
Under this instantiation, it is impossible to derive the
\C{MonotonicVis} (\S\ref{sec:ansi-isolation}) property needed to show
that successive reads in a session witness monotonically increasing
state. However, assuming {\sc cc} store consistency for reads allowed
us to complete the proof, and recover the monotonicity invariant. In
fact, we were able to complete the proof even after relaxing the
serializability requirement for increments, just by changing the
definition of the $\interp{S}(X)$ (\S\ref{sec:opsem}) to return
the value of the largest write-to-$X$ instead of the latest
write-to-$X$ in $S$.  In other words, if the reads chose the largest
valued write effect, instead of the latest write effect they witness,
then counter increments need not be serializable for reads to witness
a monotonically increasing count.\footnote{The trade-off is that the
view count is now no longer accurate, although it grows
monotonically.} We implemented this version of the counter as an
op-based RDT, which uses {\sc cc} reads and {\sc ec} increments, both
un-isolated (i.e., \emph{Read Uncommitted}), and obtained a 52\%
improvement in overall latency compared to the version that uses
serializable increments.
\vspace*{-3pt}
